---
title: 'Week 1: Scientific Method'
author: "Julie Mallon"
date: "7/31/2017"
output: 
  ioslides_presentation
   #incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Scientific Method

## Inquiry Wheel
![](filepath)

## My Research
Observation: vultures can stay aloft without using thermal soaring  
> Why would they use another type of soaring?

## Hypotheses
Alternate: They thermal and contorted soaring under different weather conditions  
Null: Use is not driven by weather conditions    

> How to test this?  
> Compare the weather conditions they used during each soaring type

## Results  
#. Vultures do not use thermal soaring when it rains    
#. Vultures do not use the new soaring type when it rains   
<br> 

What does this mean for my hypothesis?   
What are my conclusions?  

## Results
What I know: Vultures do not soar when it rains  

Observation: Vultures can stay aloft without using thermal soaring  

Hypothesis: Vultures use this new soaring type when they can’t use thermal soaring  

Results:   
1) Vultures use thermal soaring mostly when it is sunny  
2) Vultures use the new soaring type mostly when it is cloudy  
3) Vultures use both soaring types equivalently when it is partly cloudy  

> Conclusions? Next Steps?


## A Hypothesis is an explanation  
- Based on prior knowledge  
- Explanatory  
- Identifies the independent and dependent variables  
- Testable and **Falsifiable**  
- Compare to other hypotheses  
- Null HO  
- Alternate HA  


## Identify the research question, hypothesis and predictions  
#. Flashlight does not work because light bulb is burnt out  
#. If I change the light bulb, the flash light will work  
#. Why doesn’t the flashlight work?  

## What made it a hypothesis?  
- Flashlight does not work because light bulb is burnt out  
- Testable  
- Falsifiable  
- Explanatory  

## Why was the other a prediction?  
- If I change the light bulb, the flash light will work  
- Falsifiable  
- Testable  
- Not explanatory  

## Alternate versus Null Hypotheses  
**Alternate Hypothesis (HA)**    
- The independent variable has an effect on the dependent variable    
- X has an effect on Y 
  

**Null Hypothesis (H0)**   
- The independent variable has no effect on the dependent variable   
- X has no effect on Y   

## Accept or reject hypothesies
- ‘Reject’ or ‘Fail to reject’ the Null Hypothesis
- ‘Accept’ the Alternate Hypothesis
- Results ‘support’ the alternate/null hypothesis

## Proverbial Practice
You will be assigned a saying or phrase.  

With your group, write your hypotheses and predictions, and draw or write a description of a design an experiment that would test this.  

Include as much detail as possible, identifying at least independent/dependent variables, measurements, samples, etc.

  
## Proverbial Practice 
 Now, switch with the group in front/behind you. Evaluate each of the following:  
   - Hypothesis is testible and falsifiable  
   - Prediction corresponds to results  
   - Experiment tests their hypothesis  
   - Sample size is sufficient  
   - Are variables appropriate?  
  
  What improvements can you suggest?  


## Unicorn conservation project 
Scientists have recently discovered a population of unicorns has been living in the middle of a remote forest in central South America. It appears that they are at carrying capacity in their current range.  

>- The government wants to conserve this newly discovered species and it is our task to determine how to maximize the conservation effort using the scientific method.  
>- There are several plots for sale that vary in size, habitat, and distance from the current population.   
>- How can we design an experiment to maximize conservation efforts?  
>- What are our predictions for the effect of each variable on patch colonization?
>- What will we measure?

## Save da Unicorns!
- Go to terpconnect.umd.edu/~jmmallon/teaching/R.html
- Download the 'unicorn.R' file
- Open R
- Source the functions ____() and ____()
- Now input which variable you are testing to the ___ function and run the code!
- Count the number of unicorns on each island and record your data
- Using the data you collected and the plot you see in R, lets replicate this plot in excel

```{r,include=FALSE,eval=FALSE}
## Hypotheses
- Each group will test one of the variables from the board  
- Decide…  
- What is your control?  
- How can you quantify termite response?  
- How will you replicate it?  
- Create a Data Table  
- Think about your results in context of the inquiry wheel! What do you know now?  

```

## Extra question
- What is the difference between a prediction and a hypothesis?
- Give me a made up example for each! The more creative the better

- See you in 2 weeks!





